---
title: "Creating a Seurat object with multiple samples"
output: md_document
date: "2023-10-27"
---

# Building more complex Seurat objects

In most cases, we will want to go beyond reading one sample into Seurat and performing biological analyses. We often want to create a Seurat object that has multiple samples (e.g. biological replicates or patients). These samples will also have unique pieces of metadata, such as HPV- PBMC or HPV+ TIL from our [head and neck cancer dataset](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE139324).

# Prerequisites 

This analysis assumes that you have download the samples described in the GEO download tutorial under utilities. 

# Load packages 

First, we will load the necessary packages.

```{r load_packages}

library(Seurat)
library(dplyr)
library(patchwork) 

```

# Read the individual files into a list 

The data I downloaded using the tutorial are located in /home/rstudio/docker_rstudio/data/geo_download. We will create a simple for loop to read all the files into individual variables stored in a list.

```{r read_data}

full_path <- "/home/rstudio/docker_rstudio/data/geo_download/"

raw_files <- list.files(paste(full_path))
# Not the metadata... we will use this in a second 
raw_files <- raw_files[2:5]
raw_files

# Create list for Seurat object
raw_list <- vector("list",length=length(raw_files))
raw_list # empty list of length 4

# Loop to read filtered feature barcode matrices into R data into R
for (i in 1:length(raw_list)) {
  
  raw_list[[i]] <- Read10X(paste(full_path,raw_files[i],sep=""))
  
}

dim(raw_list[[1]]) # 33694 genes by  2445 cells

raw_list[[1]][1:50,1:10] # sparse matrix with counts 

# Name each item in the list by its title
names(raw_list) <- raw_files
names(raw_list)

```

# Read in metadata 

Before we construct the combined object, we will need the metadata.

We can read that in from the tsv file using readr::read_tsv (readr is a package, and the "::" lets us call a function from that package without loading the package)

```{r metadata}

# Read in the metadata
metadata_file <- readr::read_tsv("/home/rstudio/docker_rstudio/data/geo_download/GSE139324_metadata.tsv")

# Re-name a few columns for easy use 
colnames(metadata_file)[c(43,45)] <- c("disease_state","tissue")

# Filter to only our samples of interest
metadata_file_filtered <- metadata_file %>%
  filter(title %in% raw_files)

# Make sure the order of names matches the raw data list 
identical(metadata_file_filtered$title,names(raw_list))

# Create an empty list of Seurat objects
ser_list <- vector("list",length=length(raw_list))

# Loop to Create individual Seurat objects and include metadata 
for (i in 1:length(ser_list)) {
  
  # Create Seurat Object
  ser_list[[i]] <- CreateSeuratObject(raw_list[[i]])
  
  # Create metadata for each object
  meta_add <- data.frame(matrix(data=NA,nrow=ncol(ser_list[[i]]),ncol=3))
  colnames(meta_add) <- c("title","disease_state","tissue")
  rownames(meta_add) <- colnames(ser_list[[i]])
  meta_add$title <- metadata_file_filtered[i,"title"] %>% pull()
  meta_add$disease_state <- metadata_file_filtered[i,"disease_state"] %>% pull()
  meta_add$tissue <- metadata_file_filtered[i,"tissue"] %>% pull()
  
  # Add Metadata
  ser_list[[i]] <- AddMetaData(ser_list[[i]],metadata=meta_add)

}

# Check out our list of Seurat objects
ser_list

```

# Merge list of Seurat objects into a single seurat object 

```{r merge_ser}

ser_merged <- merge(ser_list[[1]],ser_list[2:4])

ser_merged@meta.data$title %>%
  table()

ser_merged@meta.data %>%
  select(title,disease_state) %>%
  table()

ser_merged@meta.data %>%
  select(title,tissue) %>%
  table()

```

# Analysis of all 4 samples 

As per the shortcut in the PBMC vignette, here's a quick analysis of the 4 healthy donor samples we merged into one Seurat object.

```{r analysis_workflow}

ser_merged <- ser_merged %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA()

ElbowPlot(ser_merged)

ser_merged <- ser_merged %>%
  FindNeighbors(.,dims=1:15) %>%
  FindClusters(.,res=c(0.3,0.5,0.7)) %>% 
  RunUMAP(.,dims=1:10)

```

# Plot a UMAP of clusters by sample

```{r plotting}

DimPlot(ser_merged,group.by="RNA_snn_res.0.3",label=T)

DimPlot(ser_merged,group.by="RNA_snn_res.0.3",split.by="title",label=T,
        ncol=2)

```

# Save output 

```{r save_results}

save_path <- "/home/rstudio/docker_rstudio/data/"
saveRDS(ser_merged,file=paste(save_path,"ser_merged_4_hd_pbmc_231027.rds",sep=""))

```
